Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections

### 2025-07-04 response_memo.qmd

---
title: "Response Memo"
format: 
  hikmah-response-typst:
    toc: false
    mainfont: TeX Gyre Pagella
    fontsize: 11pt
    bibliography: references.bib
    bibliographystyle: american-political-science-association.csl

@@

---

I thank the Reviewers for taking time to provide their thoughtful and constructive comments. I revised much of the initial manuscript by restructuring certain sections and added clarification where needed. The Reviewers' comments are cited or summarized in [red italicized text]{.memo-reviewer-inline}, and my responses are in black text. I include excerpts from my revised manuscript in [blue]{.memo-excerpt-inline}.

@@

# General Reviewer concerns on theory and competing explanations

Both reviewers' concerns touched on the same general issue, and the offered critiques and recommendations complemented each other. [*Reviewer 1*]{.memo-reviewer-inline} notes a general lack of discussion on potential explanations or explicit conjecture on a potential mechanism. [*Reviewer 2*]{.memo-reviewer-inline} recognized the same as ambiguity within of the theory; synthesis of the context (i.e., background section) and the literature didn't lead to a clear and concise reason for "why veterans" would be expected to influence confidence in elections. I dedicated a separate section titled as such hoping to clarify the source of my stated expectations, however I agree that the issue needs to be expounded at both ends.

These additions and revisions are extensive, yet I believe adding a refined theory section and discussion section addresses the most immediate concerns raised by both reviewers. The theory section is revised to make clear "why veterans", noting the gaps in available literature while making the intent of the current study explicit. The added discussion section complements the theory section by discussing potential explanations.

# Reviewer 1

@@

::: {.memo-reviewer}
***Reviewer 1***---***what is the mechanism here?***: We understand that veterans are generally viewed positively, and that’s part of why this approach might work. However, in the data, we don’t see any direct measures of how respondents feel about veterans—there’s no pre-treatment baseline for that. I understand that the data might not allow for it, or that asking upfront could interfere with the treatment, but still---it leaves us wondering: what exactly is driving the results? Is it the idea of integrity, a feeling of safety or strength, or a sense of national service? We don’t really know. For this paper, it might be acceptable to say we’re building on prior work and that the cue is sufficient, but it would be helpful to acknowledge that gap more directly and discuss potential explanations.
:::

In response to [*Reviewer 1*]{.memo-reviewer-inline}'s comment above, I included the following text in the '**Limitations**' section,

::: {.memo-excerpt}
## Limitations

The bluntness of the experimental stimulus and survey items don't allow us to establish exactly what it is about military veterans in particular that cues stronger confidence in elections. No survey items assessed public attitudes about military veterans specifically, pre- nor post-treatment. This unfortunately leaves no room for me to postulate and examine any potential mechanisms underlying results.
:::

To fully address the above comments of both reviewers, I extended a discussion prior to the conclusion on potential explanations and avenues for future research.

::: {.memo-excerpt}
## Discussion

```{=typst}
#sym.dots.h.c

```

A question lingers about whether some antecedent mechanism transmits (i.e., mediates) the effects of the veteran treatment vignette to the outcome. To put simply, this study reiterates that there *is* something special about veterans in particular that elicits greater confidence in elections, but cannot go further in determining exactly *what* that special something might be. Factors that *moderate* the relationship of the treatment on the outcome of interest are likely plenty, but one can only offer conjecture of a *mediator* driving results.

One potential explanation is that veterans, and all those who serve in the military, potentially function as a national symbol. The group is abstract just as the nation is abstract, and the potency of the veteran group emulates the potency of a national symbol. Recall Benedict Anderson's [-@anderson2006] meditation on the Tomb of the Unknown Soldier, "...saturated with ghostly *national* imaginings" [-@anderson2006, 9, emphasis original]. In various studies, national symbols such as the flag, the national anthem, or even mere appeal to national identity have potent, yet variable, effects on different outcomes [@schatz2007; @butz2007; @kemmelmeier2008; @butz2009; @gangl2016; @kalmoe2016; @levendusky2018]. It is plausible that military service members and veterans function in much the same way---as a symbol of national loyalty, pride, and nationalism for people to praise and revere out of a sense of duty prescribed by social norm and reinforced by national sentiment. As such, it may be the case that the treatment effect on confidence in elections is mediated through an affective response to a national symbol---in this case, veterans. The veteran cue may indeed stimulate national sentiment and subsequently boost confidence in elections.

@@

::: {.memo-reviewer}
***Reviewer 1***---***Hypothesis structure could be clearer***: You present H1 very clearly, which is great. However, later in the paper, there are additional expectations that seem like they could have been articulated as formal hypotheses too—perhaps H2, H3, or even sub-parts like H1a, H1b. While it’s not a major issue, clearer hypothesis structure would enhance readability and help clarify the paper’s logic as it progresses. Currently, it feels like a mix of formal and informal framing.
:::

Agreed. I revised and inserted the following hypotheses more formally, and also inserted text to make more explicit the reasoning behind the latter hypotheses. The text of the '**Hypotheses**' section is unchanged save for inclusion of the following text.

::: {.memo-excerpt}
## Hypotheses

```{=typst}
#sym.dots.h.c
```

@@

> H~4a~: Concerns about the potential for violence, threats of violence, or intimidation while voting in person will be lower among those presented with an announcement that election officials are recruiting veteran service members for election jobs compared to those presented with an announcement of recruitment efforts that do not mention military veterans.

> H~4b~: Confidence that voters will be safe to vote in-person at election sites will be higher among those presented with an announcement that election officials are recruiting veteran service members for election jobs compared to those presented with an announcement of recruitment efforts that do not mention military veterans. 
:::

::: {.memo-reviewer}
***Reviewer 1***---***Racial dynamics and trust***: This aspect also emerged for me. The sample is mostly white, making it challenging to draw conclusions about how this might work for communities of color. As we know, these communities often have long histories of institutional distrust—including towards the military. Thus, it’s possible the cue might not resonate in the same way, or perhaps it does, as many veterans are people of color too. We don’t know, and the current data cannot address that limitation—but acknowledging it and speculating could be beneficial.
:::

I included the following text in the discussion to address [**Reviewer 1**]{.memo-reviewer-inline}'s comment above.

::: {.memo-excerpt}
The data of the current study is limited in helping to discern whether the treatment garners greater confidence in elections uniformly across different racial or ethnic groups given that the largest proportion of the sample identified their race as White (77%). It is reasonable to suspect that the treatment vignette might differ by race and ethnicity, but the sample is insufficient to make such generalizations. Nevertheless, attitudes about veterans that differ by group may play an important role in determining whether publicized efforts to recruit military veterans boosts confidence in elections uniformly or whether the effect is merely partial to certain segments of the population.

In their research examining whether groups differed in attitudes about veterans, @kleykamp2018 could only speculate as to why levels of support were lowest among Hispanic and Black respondents. Their study was distinctive in that it stands as a rare case in which public attitudes about veteran military members was the specific focus of inquiry as opposed to studies that examine public opinion about the military in general. Other research has found that attitudes about the U.S. military generally---though not about active or veteran military service members in particular---differs by race and ethnicity to varying extents, but also could not supply reason for one groups' sentiment over another^[Note, however, that the sample examined by @nichols2015 was constrained to Texas residents, and the military-related issues concerned, among other things, favor for reinstatement of a military draft, support for allowing undocumented immigrants who live in the U.S. to serve in the U.S. Armed Forces, and whether undocumented immigrants should be granted U.S. citizenship as a result of military service. Although the authors found high overall support for the military as an institution, the other military-related attitudes of interest diverge far from public opinion about veterans. As such, differential effects found among different racial or ethnic groups likely have little to do with attitudes about former or active military service members.] [@nichols2015].
:::

@@

::: {.memo-reviewer}
***Reviewer 1***---***How generalizable are these results***: I appreciate your honesty about how the strongest effects appear in Maricopa County, and that we don’t observe the same impact when people consider their own local elections. That’s interesting in itself, but it raises questions—can this approach be effective in other areas? Would appointing veterans as officials in lower-salience places have any impact? Or is this cue only meaningful in scrutinized locations? You touch on this briefly, but expanding this section would add depth, particularly for readers contemplating its implications for future elections.
:::

To address these concerns, I include the following text in the '**Limitations**' section

::: {.memo-excerpt}
The setting of the vignette and specified in survey items might raise additional concerns on whether results are generalizable. Naming a specific county in the vignette adds in a potentially influential factor unaccounted for in the survey; some people may have attitudes about the county in question, others won't, while others may be muffed to consider a random county in the U.S. they know nothing about. Naming a real and somewhat salient county potentially undermines confidence that the treatment effects are solely attributable to veterans to an unknown degree. Simply put, there may be something special about Maricopa County, AZ in particular that generalizes across attitudes of trust or distrust in elections.

Arguably, however, Maricopa County is indicated as the setting in both the treatment and control vignettes, thus the only actual difference between conditions is constrained to the target group being recruited. Future research may employ vignettes naming a fictional U.S. county to hopefully eliminate the setting as a concern. Alternatively additional survey items that ask directly about one's knowledge or sentiment about the county in question could inform researchers on whether the named county imposes any undue influence at all, or whether the setting is associated with other factors such as partisanship.

As stated previously, the survey was fielded on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. The results presented here are not assumed to be generalizable to the national population. It may be feasible to weigh the sample to demographics of the population at the time the survey was fielded, but this is unnecessary for the experimental purpose of the study. If similar findings complement or replicate these results, then the expense of drawing samples more representative of the national population may be justified. At this stage, however, the results are limited to demonstrating treatment effects.

Note also that the measurement items of the components trust and distrust, and the method of computing scores for the dependent variable, confidence in elections, are unique to this study but falls under the methodological framework of Classical Test Theory (CTT). Although survey measurement items were either directly drawn from or inspired by similar survey items found in various other surveys (e.g., SPAE, Pew), the question wording, response choices, coding, and computation of sum scores of the dependent variable are specific, thus rendering measurement of the hypothetical constructs of interest (i.e., trust, distrust) less comparable across similar studies that profess measurement of similar constructs [@widaman2024]. Namely, trust, distrust, or confidence in elections as defined and measured here.
:::

# Reviewer 2

@@

::: {.memo-reviewer}
***Reviewer 2***---***Theory***: "The author could be clearer about the puzzle: Why would identifying poll workers as veterans plausibly increase confidence in elections? Why might it not? The paper gestures at answers---trust, apolitical reputation, shared identity---but doesn’t quite fully develop the competing
expectations."
:::

In response to [*Reviewer 2*]{.memo-reviewer-inline}'s comments concerning the theory, I restructure the theory section to incorporate the "Why veterans" section, expand on the premises, and add literature that better leads to the expectations. The text below reflects the revised theory section that better incorporates "Why Veterans".

::: {.memo-excerpt}
## Why Veterans

Election officials don't focus on recruiting people of a particular demographic, status, class, or prestige in order to shore up public trust in elections, and they don't discriminate recruitment efforts based on who the public thinks is best suited to secure elections from fraud.

@@

There was a sudden push to target veterans for recruitment efforts that arose shortly after the events on Capitol Hill on January 6th, 2021. After the 2020 election, large efforts were made to recruit military veterans and their families to work or volunteer as election staff [@wetheveterans2022; @nflfootballoperations2022]. Prior to that point, young people were sometimes given special mention as targets of election worker recruitment efforts [@powerthepolls2020; @herndon2020]. Generally, however, recruitment efforts cast a wide net, indiscriminate of who applies [@ross2020; @conde2020].

One can speculate that the motivation to associate military veterans with civic engagement and democracy may be intended to counter negative perceptions and impressions given by the proportion of veteran service members arrested for taking part in the events on January 6th [@milton2021; @loewenson2023; @jensen2022a]. Especially with regard to research demonstrating that willingness to support violent efforts to overturn election results (in support of Trump) is, on average, more common among veterans than among matched samples of non-veterans [@pape2024]. This is in addition to a strengthened association portrayed in media outlets between military veterans and militias [@steinhauer2020]. Prior research has substantiated such a connection between veterans and militia groups. @cooter2024 notes from her 3-year ethnographic fieldwork among Michigan militia members that, "...approximately 40% of militia leaders and 30% of members had previous military experience. Most of these veterans actively sought out such groups, as opposed to being recruited by them" [see also @cooter2013]. Thus, countering such associations by promoting a different image of veterans to the mass public and veterans alike seems like a reasonable motivation. Yet such speculation is just that.

The key findings of literature on confidence in elections hardly point to public regard for veterans as a means for improving overall confidence in elections and improving feelings of voter safety at election sites. Comparable research finds that confidence in elections is increased through elite rhetoric from fellow partisans affirming legitimacy of elections [@clayton2023], or through the provision of information about the electoral process or other official messaging of election officials themselves [@gaudette2025]. Neither of the studies just cited, however, consider attitudes or sentiment about veterans as a source for dispelling distrust and improving trust in elections. To be clear, there is no immediate "puzzle" in the literature that immediately draws out public sentiment for military veterans as a source for bolstering confidence in elections. That being said, the general public perception, attitudes, or even stereotypes about military veterans are significant to consider.

@@

::: {.memo-excerpt}
# Theory

This theory is constrained to public confidence in the administration of elections in the United States. To be clear, the scope conditions in which this theory applies concern public trust and distrust in election administration in the anticipatory period prior to the election event. Moreover, this study is, to a limited extent, indirectly related to public regard about veteran members of the U.S. Armed Forces.
:::

::: {.memo-reviewer}
***Reviewer 2***---***Theory***: Terms of art should be defined with precision. What do we mean by “voter confidence”? Is it about procedural trust? Belief in outcome integrity? Perceived impartiality? Greater clarity here will help structure the analysis and give readers clearer expectations.
:::

I agree with [**Reviewer 2**]{.memo-reviewer-inline} and apologize for any confusion. In order to address the above concern, I inserted the following text which was initially cut from an earlier draft for length. This section follows the added text above within the theory section and also further clarifies scope conditions.

@@

This means that a person's baseline level of confidence in elections must take into account their present degree of distrust---i.e., their expectations of electoral fraud. Something may cause a person to have greater expectations that election fraud will occur, which in turn will always lower their trust in elections. In contrast, a person's trust in the integrity of the electoral process is reduced by the extent to which they expect that electoral fraud will occur.

Thus, it follows that a person distrusts the integrity of the electoral process to the extent that they are to some extent certain that electoral fraud will occur or has occurred. They are, perhaps, poised to discount election results as invalid or illegitimate in advance. Either way, both trust and distrust refer to some positive degree of confidence (i.e., assurance) regarding one's expectations. They are distinguished by reference to either the normative idea about what *should* come about, or what is *supposed* to come about given the logical structure and integrity of the process. If the structure and integrity of some process is coherent and secure, respectively, then the outcome that follows will be valid. Should the integrity of that process be damaged or opaque, then the process becomes vulnerable and validity of the results are open to challenge. With respect to the administration of U.S. elections, the confident voter expects that votes will be counted as voters intend, election staff will competently administer elections, election technology will be secure from nefarious tampering, and that the process will be fair for all involved.
:::

Each reviewer raised similar concerns about the potential for differential effects of the veterans cue.

As for [*Reviewer 2's*]{.memo-reviewer-inline} concern about inclusion of a manipulation check, there were no *a priori* expectations regarding how respondents would cognitively interpret the veteran treatment vignette beyond what was stated in the hypotheses. That is, there's no reason to suspect that respondents would misinterpret or confuse "military veterans" within the treatment vignette as a reference to some particular racial group or gender, or anything other than veteran military service members.

I can't be sure as to whether respondent's views of the military generalize to attitudes about active service members and veterans, however that is a separate question that deserves further inquiry since it isn't adequately acknowledged or addressed in the literature. I included note of this with respect to discussion of the current study.

:::{.memo-excerpt}
Other studies have drawn out disparities in public attitudes about veterans but in ways that are mostly indirect, i.e., without assessing public regard for veterans directly. One overarching issue is that there's limited research that differentiates public regard for veteran and active military service members from the U.S. military as an institution in general. Some studies assess public regard for veterans indirectly by examining whether, and to what extent, endorsement of certain policies differentiate by whether such policies benefit veterans or the public generally. In one such study by @tsai2021, Republican partisanship was shown to be the most consistent factor associated with more favorable attitudes (e.g., compassion) toward homelessness and PTSD among the veteran population, but not among the general adult population. Additionally, Republican partisans tended to show greater support for federal funding for homelessness and PTSD research when the reference population was military veterans. That study, however, suggests that the effect of the veteran status cue is sensitive to the topic in question.

This comment touches on similar concerns raised by [**Reviewer 1**]{.memo-reviwer-inline} with respect to differential effects of the treatment by group. The added text in the '**Discussion**' section addresses this comment.

@@

::: {.memo-excerpt}
The median age was $46$ (mean age was $47$), $51.13\%$ ($n = 658$) women, $46.46\%$ ($n = 598$) men, and approximately $1.24\%$ who identified as either Non-binary/third gender ($n = 7$) or preferred not to say ($n = 9$). The sample primarily identified as White or Caucasian $75.76\%$ ($n = 975$), while all other non-White respondents comprised $23.08\%$ ($n = 297$) of the sample. Those who held a graduate level degree (e.g., Master's, Doctorate, or Professional level) comprised $12.9\%$ ($n = 166$) of the sample; those with either degree at the Associate or Bachelor's level comprised $35.82\%$ ($n = 461$), while $21.99\%$ ($n = 283$) had some college but no degree; and $28.13\%$ ($n = 362$) had either a high school level or equivalent education or less than high school. The largest proportion of the sample were Democrat at $43.98\%$ ($n = 566$), followed by Republicans at $41.96\%$ ($n = 540$). The proportion of true Independents^[Independent 'leaners' were grouped into the respective party affiliation in which they lean.] was $12.59\%$ ($n = 162$).

Among the sample, $34.11\%$ ($n = 439$) said they had an immediately family member who was currently serving or had previously served in the U.S. military; $64.72\%$ ($n = 833$) didn't have a family member who served. Approximately $8.24\%$ ($n = 106$) of the sample were veterans, while $1.17\%$ ($n = 15$) reported to be actively serving. 
:::

::: {.memo-reviewer}
***Reviewer 2***---***Findings***: The author identifies heterogeneous effects by election level (local vs. national), which is a fascinating and potentially important finding. But what does this mean substantively? Do voters view the stakes of “local” and “national” elections differently with respect to legitimacy? Do they associate veterans with one level more than another?
:::

### 2025-07-04 response_memo.qmd

Despite the limitations noted previously, I added the text below in response to [**Reviewer 2**]{.memo-reviewer-inline}'s concerns on the potential substantive significance of the findings. This text can be found just prior to the subsection '**2020 Election Legitimacy Beliefs**'.

::: {.memo-excerpt}
The veteran cue is potentially compensatory, which supports the second hypothesis (H~2~). That is, the gap in confidence in local elections and elections in Maricopa County is reduced by the veterans cue, therefore compensating for the otherwise lower confidence and greater insecurity in Maricopa County elections. This is analogous to studies showing that Republicans support for a Democratic candidate is compensated by a Democratic candidate's veteran status. Hypothetical Democratic veteran candidates were more palatable to Republican partisans than non-veteran Democrats [@mcdermott2015]. Similarly, Democratic veterans received higher support in Senate races, especially Democratic veteran Senate candidates whose military experience included deployment to conflict zones [@richardson2022].

The point of note is that it is *Democratic* candidates who receive *more* of a benefit that comes with veteran status^[Note that the war veteran classification provided benefits to Republican Senate candidates vote share distinct from "common" veterans---i.e., Republican Senate candidates who did not deploy to areas of conflict [@richardson2022].]; the analogy being that it is elections located elsewhere that receive the benefits of the veteran cue whereas local elections do not. Note that the results can't necessarily be read as differing between elections at the local level versus national level, but rather heterogeneous effects differ by elections at the local level and elections in Maricopa County, AZ. The compensatory effect of the veteran cue may generalize to other jurisdictions other than one's local area, or the effect may be confined to Maricopa County, AZ given that this specific county had been made especially salient in the past.
:::

@@

::: {.memo-reviewer}
***Reviewer 2***---***Findings***: The author should consider the magnitude and directionality of effects in relation to expectations. For instance: Are the observed changes in confidence substantively meaningful? Are they comparable to known effects in this literature (e.g., from partisan cues or fraud frames)?
:::

While this is understandable, I was unable to compare magnitude of effects between comparable studies due to the differences in measurements used in multiple studies examining confidence in elections. To put bluntly, the percentage point differences reported in related experimental studies are mostly incomparable due to substantial differences in measurement. I include the following text within the conclusion stating this point.

@@

That being said, this study generally emulates treatment effects of comparable studies in the literature. @gaudette2025 found positive effects of official messaging from election officials on various items measuring trust in elections and expectations of electoral fraud. Similarly, @clayton2023, confined to a sample of Republican participants, found positive influence on perceptions of 2020 election legitimacy and trust in the American electoral process from official messaging by Republican politicians reaffirming the 2020 election results.
:::

::: {.memo-reviewer}
***Reviewer 2***---***Findings***: I also encourage more discussion of null results. Even where effects are weak or absent, that tells us something important—especially if veteran cues only resonate among certain audiences or in certain electoral contexts.
:::

Although I am more than sympathetic to this view, I didn't have null results to report. However, the additions included in added the **Limitations** and **Discussion** sections speak to the concerns raised here.

{{< pagebreak >}}

### 2025-07-04 docs/_tex/index.tex
manuscript update with revisions after review

The question emerged in light of the recent surge of efforts nationwide
to recruit veterans to serve as election workers
(\citeproc{ref-lawrence2022}{Lawrence 2022};
\citeproc{ref-wentling2024}{Wentling 2024}) midst an environment where
election insecurities were heightened and intensified. Bearing that in
mind, the heightened insecurities in anticipation of upcoming elections
do not merely pertain to the election outcome (e.g., who wins), but to
one's general sense of confidence in the integrity of elections
administration and to the safety of casting a ballot in-person.

@@

@@

As stated previously, the survey was fielded on a non-probability sample
of 1,287 U.S. citizens 18 years of age or older. The results presented
here are not assumed to be generalizable to the national population. It
may be feasible to weigh the sample to demographics of the population at
the time the survey was fielded, but this is unnecessary for the
experimental purpose of the study. If similar findings complement or
replicate these results, then the expense of drawing samples more
representative of the national population may be justified. At this
stage, however, the results are limited to demonstrating treatment
effects that are sample dependent.

It is difficult to compare magnitude of effect with other studies given
that the item number, item wording, measurement, and computation of
confidence in elections is unique across most individual studies to
include this one. For instance, Gaudette et al.
(\citeproc{ref-gaudette2025}{2025}) doesn't compose multiple items into
a single composite measure, but subtracts pre-treatment values from
post-treatment values for single-item measures with Likert-style
response options. Whereas Clayton and Willer
(\citeproc{ref-clayton2023}{2023}) develops a scale referred to as
``Trust in American electoral process'' by taking the mean from three
items. The item wording of that scale is markedly similar to the items
used to compose the trust scale of this study, but the range of response
options differs considerably. Thus, while most studies examining public
confidence in elections often include the core question ``How confidence
are you that your vote {[}will be/was{]} counted as you intended in the
most recent election?'', the different ways that researchers measure and
compute variable scores on the outcome are most often, and
unfortunately, sample dependent. The study of public confidence in
elections is in desperate need of a standard method of measurement, or
even a screened and validated item bank, and this study may be seen as a
baby step to that end.

@@

The data of the current study is limited in helping to discern whether
the treatment garners greater confidence in elections uniformly across
different racial or ethnic groups given that the largest proportion of
the sample identified their race as White (77\%). It is reasonable to
suspect that the treatment vignette might differ by race and ethnicity,
but the sample is insufficient to make such generalizations.
Nevertheless, attitudes about veterans that differ by group may play an
important role in determining whether publicized efforts to recruit
military veterans boosts confidence in elections uniformly or whether
the effect is merely partial to certain segments of the population. In
their research examining whether groups differed in attitudes about
veterans, Kleykamp, Hipes, and MacLean
(\citeproc{ref-kleykamp2018}{2018}) could only speculate as to why
levels of support were lowest among Hispanic and Black respondents.
Other research has found that attitudes about the U.S. military
generally---though not about active or veteran military service members
in particular---differs by race and ethnicity to varying extents, but
could not supply reason for one groups' sentiment over another
(\citeproc{ref-nichols2015}{Nichols 2015}).

Other studies have drawn out disparities in public attitudes about
veterans but in ways that are mostly indirect, i.e., without assessing
public regard for veterans directly. One overarching issue is that
there's limited research that differentiates public regard for veteran
and active military service members from the U.S. military as an
institution in general. For instance, some studies assess public regard
for veterans indirectly by examining whether, and to what extent,
endorsement of certain policies differentiate by whether such policies
benefit veterans or the public generally. In one such study, Republican
partisanship was shown to be the most consistent factor associated with
more favorable attitudes on homelessness and PTSD among the veteran
population, but not among the general adult population
(\citeproc{ref-tsai2021}{Tsai et al. 2021}).

The bluntness of the experimental stimulus and survey items don't allow
us to establish exactly what it is about military veterans in particular
that cues stronger confidence in elections. No survey items assessed
public attitudes about military veterans specifically, pre- nor
post-treatment. This unfortunately leaves no room for me to postulate
and examine any potential mechanisms underlying results. As such, a
question lingers about whether some antecedent mechanism transmits
(i.e., mediates) the effects of the veteran treatment vignette to the
outcome. To put simply, this study reiterates that there \emph{is}
something special about veterans in particular that elicits greater
confidence in elections, but cannot go further in determining exactly
\emph{what} that special something might be.

Factors that \emph{moderate} the relationship of the treatment on the
outcome of interest are likely plenty, but one can only offer conjecture
of a \emph{mediator} driving results. One potential explanation is that
veterans, and all those who serve in the military, potentially function
as a national symbol. The group is abstract just as the nation is
abstract, and the potency of the veteran group emulates the potency of a
national symbol. Recall Benedict Anderson's
(\citeproc{ref-anderson2006}{2006}) meditation on the Tomb of the
Unknown Soldier, ``\ldots saturated with ghostly \emph{national}
imaginings'' (\citeproc{ref-anderson2006}{2006, 9}, emphasis original).
In various studies, national symbols such as the flag, the national
anthem, or even mere appeal to national identity have potent, yet
variable, effects on different outcomes (\citeproc{ref-butz2009}{Butz
2009}; \citeproc{ref-butz2007}{Butz, Plant, and Doerr 2007};
\citeproc{ref-gangl2016}{Gangl, Torgler, and Kirchler 2016};
\citeproc{ref-kalmoe2016}{Kalmoe and Gross 2016};
\citeproc{ref-kemmelmeier2008}{Kemmelmeier and Winter 2008};
\citeproc{ref-levendusky2018}{M. S. Levendusky 2018};
\citeproc{ref-schatz2007}{Schatz and Lavine 2007}). It is plausible that
military service members and veterans function in much the same way---as
a symbol of national loyalty, pride, and nationalism for people to
praise and revere out of a sense of duty prescribed by social norm and
reinforced by national sentiment. As such, it may be the case that the
treatment effect on confidence in elections is mediated through an
affective response to a national symbol---in this case, veterans. The
veteran cue may indeed stimulate national sentiment and subsequently
boost confidence in elections.

@@

>>>>

@@

\subsection{Limitations}\label{limitations}

The bluntness of the experimental stimulus and survey items don't allow
us to establish exactly what it is about military veterans in particular
that cues stronger confidence in elections. No survey items assessed
public attitudes about military veterans specifically, pre- nor
post-treatment. This unfortunately leaves no room for me to postulate
and examine any potential mechanisms underlying results.

@@

@@

As stated previously, the survey was fielded on a non-probability sample
of 1,287 U.S. citizens 18 years of age or older. The results presented
here are not assumed to be generalizable to the national population. It
may be feasible to weigh the sample to demographics of the population at
the time the survey was fielded, but this is unnecessary for the
experimental purpose of the study. If similar findings complement or
replicate these results, then the expense of drawing samples more
representative of the national population may be justified. At this
stage, however, the results are limited to demonstrating treatment
effects.

The research question emerged in light of the recent surge of efforts
nationwide to recruit veterans to serve as election workers
(\citeproc{ref-lawrence2022}{Lawrence 2022};
\citeproc{ref-wentling2024}{Wentling 2024}) midst an environment where
election insecurities were heightened and intensified. Heightened
insecurities in anticipation of upcoming elections do not merely pertain
to the election outcome (e.g., who wins), but to one's general sense of
confidence in the administration of elections, the integrity of the
electoral process, and to the safety of casting a ballot in-person. This
study demonstrated that information neutral to the trustworthiness of
election administration can not only improve trust in elections, but
simultaneously reduce distrust in elections. Those who were presented
information about veteran recruitment efforts reported greater trust
that votes will be counted accurately, that election staff will
competently administer elections, that election technology will be
secure from nefarious tampering, and that the process will be fair for
all involved. Moreover, those presented with information about veteran
recruitment efforts reported lower expectations that electoral fraud
will occur. They also felt that election sites would be safe for voters
to cast a ballot in person and were less concerned about potential
violence compared to those who were not presented information about
veteran recruitment efforts.

Confidence in elections was greater among those who read the treatment
vignette announcing efforts to recruit veterans to work as election
staff and volunteers, but this effect was conditional on whether survey
questions asked about elections Maricopa County, AZ. When survey
questions asked about elections in one's local area, confidence in
elections among the treatment group was not significantly different from
the control group. However, the sample's confidence in their local area
elections was high in general relative to elections in Maricopa County,
AZ (Figure~\ref{fig-dist}). The results presented here suggest that the
effect of the veteran cue in the treatment vignette is compensatory.
That is, the gap in confidence in local elections and elections in
Maricopa County is reduced by the veterans cue, therefore compensating
for the otherwise lower confidence and greater insecurity in Maricopa
County elections.

A question lingers about whether some antecedent mechanism transmits
(i.e., mediates) the effects of the veteran treatment vignette to the
outcome. To put simply, this study reiterates that there \emph{is}
something special about veterans in particular that elicits greater
confidence in elections, but cannot go further in determining exactly
\emph{what} that special something might be. Factors that
\emph{moderate} the relationship of the treatment on the outcome of
interest are likely plenty, but one can only offer conjecture of a
\emph{mediator} driving results.

One potential explanation is that veterans, and all those who serve in
the military, potentially function as a national symbol. The group is
abstract just as the nation is abstract, and the potency of the veteran
group emulates the potency of a national symbol. Recall Benedict
Anderson's (\citeproc{ref-anderson2006}{2006}) meditation on the Tomb of
the Unknown Soldier, ``\ldots saturated with ghostly \emph{national}
imaginings'' (\citeproc{ref-anderson2006}{2006, 9}, emphasis original).
In various studies, national symbols such as the flag, the national
anthem, or even mere appeal to national identity have potent, yet
variable, effects on different outcomes (\citeproc{ref-butz2009}{Butz
2009}; \citeproc{ref-butz2007}{Butz, Plant, and Doerr 2007};
\citeproc{ref-gangl2016}{Gangl, Torgler, and Kirchler 2016};
\citeproc{ref-kalmoe2016}{Kalmoe and Gross 2016};
\citeproc{ref-kemmelmeier2008}{Kemmelmeier and Winter 2008};
\citeproc{ref-levendusky2018}{M. S. Levendusky 2018};
\citeproc{ref-schatz2007}{Schatz and Lavine 2007}). It is plausible that
military service members and veterans function in much the same way---as
a symbol of national loyalty, pride, and nationalism for people to
praise and revere out of a sense of duty prescribed by social norm and
reinforced by national sentiment. As such, it may be the case that the
treatment effect on confidence in elections is mediated through an
affective response to a national symbol---in this case, veterans. The
veteran cue may indeed stimulate national sentiment and subsequently
boost confidence in elections.

@@

It is difficult to compare magnitude of effect with other studies given
that the item number, item wording, measurement, and computation of
confidence in elections is unique across most individual studies to
include this one. For instance, Gaudette et al.
(\citeproc{ref-gaudette2025}{2025}) doesn't compose multiple items into
a single composite measure, but subtracts pre-treatment values from
post-treatment values for single-item measures with Likert-style
response options. Whereas Clayton and Willer
(\citeproc{ref-clayton2023}{2023}) develops a scale referred to as
``Trust in American electoral process'' by taking the mean from three
items. The item wording of that scale is markedly similar to the items
used to compose the trust scale of this study, but the range of response
options differs considerably. Thus, while most studies examining public
confidence in elections often include the core question ``How confidence
are you that your vote {[}will be/was{]} counted as you intended in the
most recent election?'', the different ways that researchers measure and
compute variable scores on the outcome most often produces results that
are, unfortunately, sample dependent. The study of public confidence in
elections is in desperate need of a standard method of measurement, or
even a screened and validated item bank, and this study may be seen as a
baby step to that end.

The data of the current study is limited in helping to discern whether
the treatment garners greater confidence in elections uniformly across
different racial or ethnic groups given that the largest proportion of
the sample identified their race as White (77\%). It is reasonable to
suspect that the treatment vignette might differ by race and ethnicity,
but the sample is insufficient to make such generalizations.
Nevertheless, attitudes about veterans that differ by group may play an
important role in determining whether publicized efforts to recruit
military veterans boosts confidence in elections uniformly or whether
the effect is merely partial to certain segments of the population.

In their research examining whether groups differed in attitudes about
veterans, Kleykamp, Hipes, and MacLean
(\citeproc{ref-kleykamp2018}{2018}) could only speculate as to why
levels of support were lowest among Hispanic and Black respondents.
Their study was distinctive in that it stands as a rare case in which
public attitudes about veteran military members was the specific focus
of inquiry as opposed to studies that examine public opinion about the
military in general. Other research has found that attitudes about the
U.S. military generally---though not about active or veteran military
service members in particular---differs by race and ethnicity to varying
extents, but also could not supply reason for one groups' sentiment over
another\footnote{Note, however, that the sample examined by Nichols
  (\citeproc{ref-nichols2015}{2015}) was constrained to Texas residents,
  and the military-related issues concerned, among other things, favor
  for reinstatement of a military draft, support for allowing
  undocumented immigrants who live in the U.S. to serve in the U.S.
  Armed Forces, and whether undocumented immigrants should be granted
  U.S. citizenship as a result of military service. Although the authors
  found high overall support for the military as an institution, the
  other military-related attitudes of interest diverge far from public
  opinion about veterans. As such, differential effects found among
  different racial or ethnic groups likely have little to do with
  attitudes about former or active military service members.}
(\citeproc{ref-nichols2015}{Nichols 2015}).

Other studies have drawn out disparities in public attitudes about
veterans but in ways that are mostly indirect, i.e., without assessing
public regard for veterans directly. One overarching issue is that
there's limited research that differentiates public regard for veteran
and active military service members from the U.S. military as an
institution in general. Some studies assess public regard for veterans
indirectly by examining whether, and to what extent, endorsement of
certain policies differentiate by whether such policies benefit veterans
or the public generally. In one such study by Tsai et al.
(\citeproc{ref-tsai2021}{2021}), Republican partisanship was shown to be
the most consistent factor associated with more favorable attitudes
(e.g., compassion) toward homelessness and PTSD among the veteran
population, but not among the general adult population. Additionally,
Republican partisans tended to show greater support for federal funding
for homelessness and PTSD research when the reference population was
military veterans. That study, however, suggests that the effect of the
veteran status cue is sensitive to the topic in question.
